---
title: "Geospatial Assignment"
author: "Sarah Mixon"
date: "5/27/2018"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE)
```  


```{r packages }
library(tidycensus)
library(tidyverse)
options(tigris_use_cache = TRUE)
library(ggplot2)
library(sf)
library(tmap)
library(viridis)
library(devtools)
library(leaflet)
library(stringr)
library(DT)


census_api_key("2cb24b255d5b7f9ce4bc9f2ea6d0be84276786ff")
```  

## Demographics of Michigan Population, by County

### Population in Michigan

#### The majority of the population live in the southern portion of the state. Here we see that Detroit, and its surrounding areas, are home to a large portion of the population. 

```{r population}

# population variable
mi_pop <- get_acs(geography = "county",
                  variables = c(population = "B01003_001"),
                  state = "MI",
                  # and shapefile
                  geometry = TRUE)

# population percentiles
pal <- colorQuantile(palette = "viridis", domain = mi_pop$estimate, n = 10 )

# map of population
mi_pop %>%
  st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
              stroke = FALSE,
              smoothFactor = 0,
              fillOpacity = 0.5,
              color = ~ pal(estimate)) %>%
  addLegend("bottomright",
            pal = pal,
            values = ~ estimate,
            title = "Population Percentiles, by County",
            opacity = 1)

```  

### Income

#### The median income level is pretty consistent across the state, but we do see high levels in the southeast region of the state, around Detroit. 

```{r median income}

# median income variable
mi_med_income <- get_acs(geography = "county",
                         variables = c(medincome = "B19013_001"),
                         summary_var = "B01003_001", 
                         state = "MI",
                         geometry = TRUE)

# map of median income
pal_income <- colorNumeric(palette = "plasma",
                           domain = mi_med_income$estimate)

mi_med_income %>%
  st_transform(crs = "+init=epsg:4326") %>%
  leaflet(width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
              stroke = FALSE,
              smoothFactor = 0,
              fillOpacity = 0.5,
              color = ~pal_income(estimate)) %>%
  addLegend("bottomright",
            pal = pal_income,
            values = ~ estimate,
            title = "Median Household Income, by County",
            opacity = 1)
```  

### Race

#### The population is overwhelming white across the state. However, in the southeast region, and especially around Detroit, there seems to be an equal mix of white and black in the population. 

```{r race in michigan}
# race variables
race_vars <- c(white = "B02001_002",
               black = "B02001_003",
               am_indian = "B02001_004",
               asian = "B02001_005",
               hispanic = "B03002_012")

mi_race <- get_acs(geography = "county",
                   variables = race_vars,
                   summary_var = "B01003_001", 
                   state = "MI",
                   geometry = TRUE)
# by race by county
mi_race %>%
  # percent total of each column
  mutate(pct = 100 * (estimate / summary_est)) %>%
  # facet maps
  ggplot(aes(fill = pct, color = pct)) +
  facet_wrap(~variable) +
  geom_sf() + 
    coord_sf(crs = 26915) + 
  scale_fill_viridis() +
  scale_color_viridis() +
    labs(title = "Race by County",
       subtitle = "U.S. Census Bureau, ACS 2016")

```  

### Unemployment

#### We see relatively high rates of unemployment among those individuals with highschool and some college education.

```{r unemployment}

unemployment_vars <- c(bachelors = "B23006_028", 
                       some_college = "B23006_021", 
                       highschool = "B23006_014", 
                       less_hs = "B23006_007")

mi_unemployment_education <- get_acs(geography = "county",       variables = unemployment_vars,        state = "MI",                        summary_var = "B01003_001",              geometry = TRUE)

# unemployment by education by county
mi_unemployment_education %>%
  # percent total of each column
  mutate(pct = 100 * (estimate / summary_est)) %>%
  # facet maps
  ggplot(aes(fill = pct, color = pct)) +
  facet_wrap(~variable) +
  geom_sf() + 
    coord_sf(crs = 26915) + 
  scale_fill_viridis() +
  scale_color_viridis() +
  labs(title = "Unemployment by Educational Attainment, by County",
       subtitle = "U.S. Census Bureau, ACS 2016")
```  

##### _Data: US Census Bureau, ACS 2016_







